---
title: "Another attempt at generating data"
author: "Gustavo Garcia-Franceschini"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nph)
```

# Treatment Assignment

Here, I assign treatment, generate from uniform distribution and define parameters. Later parts of the project could explore different parameters.

```{r}
# 0 = control; 1 = treat
treat = c(rep(0, 500), rep(1,500))

#random U
u = runif(1000)

#treatment effect
beta = -0.5

#coefficients
lambda = 2
alpha = 3
```

# Proportional Data

Here, we produce our data from the proportional assumption. Notice the only difference in the distributions of the two groups is the factor of 1/exp(beta).

```{r}
proportional = tibble(
  
  treat = treat,
  
  u = u,
  
  t = 1/(lambda^(1/alpha)) * (-log(1-u)/exp(beta*treat))^(1/alpha),
  
  event = T
)

```

Here, we can see the log-rank test correctly rejects the Null (that the two groups come from the same distribution.)

```{r}
logrank.test(proportional$t, proportional$event, proportional$treat, 
             rho = 0, gamma = 0 )
```

```{r}
s = survival::survfit(survival::Surv(t, event) ~ treat, data = proportional)

plot(s)
```

```{r}
treat_f <- as.factor(c(rep("0", 500), rep("1", 500)))

df <- data.frame(surv = s$surv, treat = treat_f, time = s$time)
```

```{r}
ggplot(data = df, aes(x = time, y = surv)) + 
    geom_point(aes(colour = treat)) + 
    geom_line(aes(colour = treat)) +
    theme_bw()
```



# Effect increases

Here, both treatment groups are sampled from the same distribution, up until time t_0. If the individual is alive by time t, the treatment activates (if they were assigned treatment). Future research can look at different values of t_0.

```{r}
t_0 = 1.05

increases = tibble(
  
  treat = treat,
  
  u = u,
  
  t = 1/(lambda^(1/alpha)) * (-log(1-u)/exp(beta*0))^(1/alpha), #force treat=0
  
  event = T
)
```

```{r}
increases = increases %>%
  mutate(
    t = case_when(
      
      t < t_0 ~ t,
      
      t >= t_0 ~ 1/(lambda^(1/alpha)) * (-log(1-u)/exp(beta*treat))^(1/alpha)
      
    )
  )
```

Here, we can see that the log-rank test with the proportional assumption incorrectly fails to reject the Null.

```{r}
logrank.test(increases$t, increases$event, increases$treat, 
             rho = 0, gamma = 0 )
```

```{r} 
s = survival::survfit(survival::Surv(t, event) ~ treat, data = increases)

plot(s)
```


```{r}
df <- data.frame(surv = s$surv, treat = treat_f, time = s$time)
ggplot(data = df, aes(x = time, y = surv)) + 
    geom_point(aes(colour = treat)) + 
    geom_line(aes(colour = treat)) +
    theme_bw()
```

When adjusting the log-rank test for delayed effects, we do reject the Null.

```{r}
logrank.test(increases$t, increases$event, increases$treat, 
             rho = 0, gamma = 1 )
```

#Decreases

Here, we first sample using the treatment effect. If the individual reaches time t_0, the effect diminishes. Further research can look at changing t_0.

```{r}
t_0 = 0.8
decreases = tibble(
  
  treat = treat,
  
  u = u,
  
  t = 1/(lambda^(1/alpha)) * (-log(1-u)/exp(beta*treat))^(1/alpha),
  
  event = T
)
```

```{r}
decreases = decreases %>%
  mutate(
    t = case_when(
      
      t < t_0 ~ t,
      
      t >= t_0 ~ 1/(lambda^(1/alpha)) * (-log(1-u)/exp(beta*0))^(1/alpha)
      
    )
  )
```

Here, we can see that the log-rank test with the proportional assumption incorrectly fails to reject the Null.

```{r}
logrank.test(decreases$t, decreases$event, decreases$treat, 
             rho = 0, gamma = 0 )
```

```{r} 
s = survival::survfit(survival::Surv(t, event) ~ treat, data = decreases)

plot(s)
```


```{r}
df <- data.frame(surv = s$surv, treat = treat_f, time = s$time)
ggplot(data = df, aes(x = time, y = surv)) + 
    geom_point(aes(colour = treat)) + 
    geom_line(aes(colour = treat)) +
    theme_bw()
```

When adjusting the log-rank test for diminishing effects, we do reject the Null.

```{r}
logrank.test(decreases$t, decreases$event, decreases$treat, 
             rho = 1, gamma = 0 )
```


