---
title: "Third attempt at generating data"
author: "Gustavo Garcia-Franceschini"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nph)
library(survminer)
```

# Treatment Assignment

Here, I assign treatment, generate from uniform distribution and define parameters. Later parts of the project could explore different parameters.

```{r}
# 0 = control; 1 = treat
treat = c(rep(0, 500), rep(1,500))

set.seed(42)
#random U
u = runif(1000)

#treatment effect
beta = -0.5

#coefficients
lambda = 2
alpha = 3
```

# Proportional Data

Here, we produce our data from the proportional assumption. Notice the only difference in the distributions of the two groups is the factor of 1/exp(beta).

```{r}
proportional = tibble(
  
  treat = treat,
  
  u = u,
  
  t = 1/(lambda^(1/alpha)) * (-log(1-u)/exp(beta*treat))^(1/alpha),
  
  event = T
)

```

Here, we can see the log-rank test correctly rejects the Null (that the two groups come from the same distribution.)

```{r}
logrank.test(proportional$t, proportional$event, proportional$treat, 
             rho = 0, gamma = 0 )

logrank.test(proportional$t, proportional$event, proportional$treat, 
             rho = 0, gamma = 1 )

logrank.test(proportional$t, proportional$event, proportional$treat, 
             rho = 1, gamma = 0 )

logrank.maxtest(proportional$t, proportional$event, proportional$treat)
```

```{r}
s = survival::survfit(survival::Surv(t, event) ~ treat, data = proportional)
```

```{r}
ggsurvplot(s, conf.int = F, pval = FALSE, risk.table = FALSE,
    legend.labs=c("Control", "Treatment"), 
    legend = c(0.15, 0.15),
    legend.title = "Treatment Group",
    palette=c("#48ACF0", "#EA3546"), 
    xlab = "Time",
    ylab = "Proportion surviving",
    title = "Proportional effect model",
    subtitle = "Beta = -0.5, n = 1000",)
```




# Effect increases

Here, both treatment groups are sampled from the same distribution, up until time t_0. If the individual is alive by time t, the treatment activates (if they were assigned treatment). Future research can look at different values of t_0.

```{r}
t_0 = 1.05

increases = tibble(
  
  treat = treat,
  
  u = u,
  
  t = 1/(lambda^(1/alpha)) * (-log(1-u)/exp(beta*0))^(1/alpha), #force treat=0
  
  event = T
)
```

```{r}
increases = increases %>%
  mutate(
    t = case_when(
      
      t < t_0 ~ t,
      
      t >= t_0 ~ 1/(lambda^(1/alpha)) * (-log(1-u)/exp(beta*treat))^(1/alpha)
      
    )
  )
```

Here, we can see that the log-rank test with the proportional assumption incorrectly fails to reject the Null.

```{r}
logrank.test(increases$t, increases$event, increases$treat, 
             rho = 0, gamma = 0 )

logrank.test(increases$t, increases$event, increases$treat, 
             rho = 0, gamma = 1 )

logrank.test(increases$t, increases$event, increases$treat, 
             rho = 1, gamma = 0)

logrank.maxtest(increases$t, increases$event, increases$treat)
```

```{r} 
s = survival::survfit(survival::Surv(t, event) ~ treat, data = increases)
```


```{r}
ggsurvplot(s, conf.int = F, pval = FALSE, risk.table = FALSE,
    legend.labs=c("Control", "Treatment"), 
    legend = c(0.15, 0.15),
    legend.title = "Treatment Group",
    palette=c("#48ACF0", "#EA3546"), 
    xlab = "Time",
    ylab = "Proportion surviving",
    title = "Increasing effect model",
    subtitle = "Beta = -0.5, n = 1000, t_0 = 1.05")
```

#Decreases

Here, we first sample using the treatment effect. If the individual reaches time t_0, the effect diminishes. Further research can look at changing t_0.

```{r}
t_0 = 0.8
decreases = tibble(
  
  treat = treat,
  
  u = u,
  
  t = 1/(lambda^(1/alpha)) * (-log(1-u)/exp(beta*treat))^(1/alpha),
  
  event = T
)
```

```{r}
decreases = decreases %>%
  mutate(
    t = case_when(
      
      t < t_0 ~ t,
      
      t >= t_0 ~ 1/(lambda^(1/alpha)) * (-log(1-u)/exp(beta*0))^(1/alpha)
      
    )
  )
```

Here, we can see that the log-rank test with the proportional assumption incorrectly fails to reject the Null.

```{r}
logrank.test(decreases$t, decreases$event, decreases$treat, 
             rho = 0, gamma = 0 )

logrank.test(decreases$t, decreases$event, decreases$treat, 
             rho = 1, gamma = 0 )

logrank.test(decreases$t, decreases$event, decreases$treat, 
             rho = 0, gamma = 0 )

logrank.maxtest(decreases$t, decreases$event, decreases$treat)
```

```{r} 
s = survival::survfit(survival::Surv(t, event) ~ treat, data = decreases)
```


```{r}
ggsurvplot(s, conf.int = F, pval = FALSE, risk.table = FALSE,
    legend.labs=c("Control", "Treatment"), 
    legend = c(0.15, 0.15),
    legend.title = "Treatment Group",
    palette=c("#48ACF0", "#EA3546"), 
    xlab = "Time",
    ylab = "Proportion surviving",
    title = "Diminishing effect model",
    subtitle = "Beta = -0.5, n = 1000, t_0 = 0.8")
```


Type I Error and Power in 1000 simulations

See if results change with varying sample sizes

Try different t_0

This project assumes no censuring or truncation